A benchmark for the implementation of standard vehicle routing algorithms

### 2021-04-15 paper/main.tex
more description and comments

Purposes: determine which language we can tell our students to
use to implement routing optimisation routines. Also determine which
language we can use ourselves.

@@

\section{Experiments}
All programs are run concurrently on an AMD Ryzen 5 3600 6-Core
Processor at 3.6 GHz with 16 GB RAM running Linux. Twelve threads can
run in parallel, however in our setting at most 5 threads are used at
any given time. For every benchmark and every instance, CPU times are
reported as a ratio of a \emph{reference time} for that benchmark and
instance. In practice, for each benchmark and instance, the reference
time is the time required by the c++98 implementation for that
benchmark and that instance. Therefore the c++98 implementation always
has a CPU ratio of 1.

@@

\subsection{Impact of using a flat matrix}
\subsection{Impact of using static arrays}
\subsection{A comparison of Python implementations}
\subsection{A comparison of C++ implementations}
Also add a clang++ vs g++ comparison?
\subsection{General cross-language comparison}
\subsection{Cross-language comparison: fast languages}

\subsection{Cross-platform comparison}
We now look at the same benchmarks, using the same languages, but running
in a different environment. We compare runs on an Intel i5-8265U
4-core CPU at 1.6 GHz and on the previously mentioned AMD CPU. The
goal is not to observe which CPU provides the best performance, rather
to determine whether the experimental observations made with one CPU
are valid with the other. For this reason, we stick to reporting CPU
times as a ratio to the CPU time required by the c++98 implementation.

Perhaps also run on WU cluster and observe differences.

\section{Discussion}
The question of what programming language to use when implementing
optimization algorithms is one that, by essence, does not have a final
answer.

@@

\section{Summary of language and compiler versions}
\label{app:lang}
\begin{table}
  \begin{tabular}{l|l}
    \toprule
    Language/tool & language or compiler version & options \\
    \midrule
    C++14 & g++ & -Wall -ansi -pedantic -O3 -std=c++14 \\
    C++98 & g++ & -Wall -ansi -pedantic -O3 -std=c++98 \\
    Python & & \\
    Pypy & & \\
    Numpy & & \\
    Numba & & \\
    Java &  & \\
    Julia & 1.5.3 & --check-bounds=no --inline=yes -O3 -t 1 \\
    Rust & & \\
    JavaScript & Node & \\
    \bottomrule   
  \end{tabular}
  \caption{Overview of all implementations}
\end{table}

>>>>

Purposes: determine which language we can tell our students to
use to implement routing optimization routines. Also determine which
language we can use ourselves.

@@

\section{Experiments}
All programs are run concurrently on an desktop computer with an AMD
Ryzen 5 3600 6-Core Processor at 3.6 GHz with 16 GB RAM, running
Linux. Twelve threads can
run in parallel, however in our setting at most 5 threads are used at
any given time. For every benchmark and every instance, CPU times are
reported as a ratio of a \emph{reference time} for that benchmark and
instance. In practice, for each benchmark and instance, the reference
time is the time required by the c++98 implementation for that
benchmark and that instance. Therefore the c++98 implementation always
has a CPU ratio of 1.

@@

\subsection{A comparison of Python implementations}

\subsection{Impact of using a flat matrix}

\subsection{Impact of using static arrays}

\subsection{A comparison of C++ implementations}
Also add a clang++ vs g++ comparison?

\subsection{General cross-language comparison}

\subsection{Cross-language comparison: fast languages}

\subsection{Cross-platform comparison}
We now look at the same benchmarks, using the same languages, but running
in a different environment. We look at runs on an laptop Intel CPU
(Intel i5-8265U, 4 cores, 1.6 GHz), on a Raspberry Pi 4 (ARMv7,
4 cores, 1.5 GHz), on a shared cluster (Intel Xeon E5-2650 v3, 20
cores, 3 GHz) and on the previously mentioned AMD CPU. The
goal is not to observe which CPU provides the best performance, rather
to determine whether the experimental observations made with one CPU
are valid with the other. For this reason, we only report results that
are different from the ones observed with the previous CPU.
We note here that despite our best effort, language and compiler
versions might vary between platforms, due to the diversity in
package version availability on different systems. Ultimately, the
goal of this analysis is not to obtain a very precise assessment of
relative performance depending on compiler version and architecture, as it
is of limited use to anyone using a slightly different architecture or
compiler version; rather, we want to assess the reliability of
previously observed results.

\section{Discussion}
The question of what programming language to use when implementing
optimization algorithms is one that, by essence, does not have a final
answer. This being said, there is value in knowing how much
performance is gained or lost when using a certain
language on a certain architecture, or how much can be
gained by using different data structures. The benchmark suite
introduced in this paper is free software and can be used for that
purpose. We encourage students and researchers in
transport optimization to use it to support their implementation
decisions. Further experimental analyses similar to the ones conducted
here can also be good practice, for instance simply comparing the
runtime performance of the same C++ code when compiled with different
compilers, which is easier than it ever was.

Additionally, some insight can already be taken straight out of the
experiments conducted above. Using languages like Julia or Rust
does not incur a significant runtime penalty when compared to more
established languages like C++, and can actually be faster
sometimes. Interpreted languages like Python or JavaScript nowadays
have reasonably good runtime performance, under five times slower than
pre-compiled languages, under the right conditions (using a JIT
compiler). This means that the flexibility and ease of use of these
languages can be worth the performance hit, depending on the
application. Additionally, using JavaScript opens perspectives, as the
integration of optimization algorithms in web applications is easier
than it ever was, and can now be run client-side.

@@

\section{Summary of language and compiler versions}
\label{app:lang}
\begin{table}
  \begin{tabular}{l|l}
    \toprule
    Language/tool & language or compiler version & options \\
    \midrule
    C++14 & g++ 10.2.1 & -Wall -ansi -pedantic -O3 -std=c++14 \\
    C++98 & g++ 10.2.1 & -Wall -ansi -pedantic -O3 -std=c++98 \\
    CPython & 3.9.2 & \\
    Pypy & 7.3.3 & \\
    % Numpy & & \\
    % Numba & 0.53.1 & \\
    Java & OpenJDK version 11 & \\
    Julia & 1.6.0 & --check-bounds=no --inline=yes -O3 -t 1 \\
    Rust & 1.51 & cargo build --release \\
    JavaScript & Node & \\
    \bottomrule   
  \end{tabular}
  \caption{Overview of all implementations}
\end{table}

There are small variations in versions used on different
platforms. Most notably, the Julia version used on the ARMv7 CPU is
1.3, as it is the latest version available for that architecture.
